Manuscript draft

### 2024-04-17 manuscript.rmd
switched to rmd so I can try using trackdown/google docs

@@

```{r}
#| include: false
library(tidyverse)
library(here)
tar_config_set(store = here::here('_targets'))
```

@@

```{r}
tar_load(cleaned_r)
tar_load(joined0_r)
pctremovedcleaned <- (joined0_r[1]-cleaned_r[1])/joined0_r[1]
```

@@

```{r}
tar_load(downsampled_10min_forSocial_r)
nvultures <- map_dbl(downsampled_10min_forSocial_r, "vultures")
```

@@

```{r}
tar_load(downsampled_10min_r)
nvultures <- map_dbl(downsampled_10min_r, "vultures")
```

After removing days with too few fixes per day, we removed any vultures that did not have at least 30 days of valid movement data remaining in a given season, to ensure that movement estimates and averages were based on a high enough sample size of days to capture the range of the vulture's behavior. Since this was done on a per-season basis, individuals that were removed from one season's analysis due to not having enough valid days could still have movement measures calculated for a different season. Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes. Ultimately, we calculated movement measures for a total of `r nvulturesmov` **focal individuals** in total (range: `r min(nvulturesmov_szns)`-`r max(nvulturesmov_szns)`; mean = `r round(mean(nvulturesmov_szns), 2)` vultures per season, sd = `r round(sd(nvulturesmov_szns), 2)` vultures per season).

@@

```{r}
#| include: false
load(here("data/calcMovement/pca_space.Rda"))
propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]

load(here("data/calcMovement/pca_roost.Rda"))
propvar_roost <- as.data.frame(summary(pca_roost)$importance)$PC1[2]

load(here("data/calcMovement/pca_movement.Rda"))
propvar_movement <- as.data.frame(summary(pca_movement)$importance)$PC1[2]
```

@@

**Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity. It explained `r 100*round(propvar_movement,2)`% of the variation in the PCA. Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall. **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE. It explained `r 100*round(propvar_space,2)`% of the variation in the PCA. Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e. the core area made up a larger proportion of the home range). **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night). It explained `r 100*round(propvar_roost,2)`% of the variation in the PCA.

@@

```{r}
#| include: false
#| output: false
source(here("scripts/04.0_mixedModels.R"))
```

@@

```{r}
#| eval: false
#| output: false
#| label: fig-d-space
#| fig-cap: "Relationship between normalized degree and space use"
# NOT SHOW
plot_d_eff_space # null--no overall relationship between space use and degree
```

```{r}
#| label: fig-d-space-situ
#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
plot_d_eff_situ_space
plot_d_emt_situ_space
```

There was a significant relationship between how vultures used space and how many other individuals they encountered in flight. Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others. There was a slight, albeit non-statistically-significant, trend toward lower degree in the co-feeding and co-roosting networks for vultures with larger ranges.

```{r}
#| eval: false
#| output: false
#| label: fig-d-roost
#| fig-cap: "Relationship between normalized degree and roost diversification"
# NOT SHOW
plot_d_eff_roost # null--no overall relationship between roost diversification and degree
```

```{r}
#| eval: false
#| output: false
#| label: fig-d-movement-situ
#| fig-cap: "Relationship between normalized degree and movement, with separate trend lines for each social situation"
# NOT SHOW
plot_d_eff_situ_movement
plot_d_emt_situ_movement # going to show the plot by season instead of this one
```

```{r}
#| label: fig-d-movement-situ-season
#| fig-cap: "Relationship between normalized degree and movement, for each season and social situation."
plot_d_emt_situ_movement_season
```

Movement behavior affected degree in all three social situations. Vultures with a higher movement score co-roosted with fewer conspecifics only in the transient, pre-breeding season; otherwise movement did not affect their co-roosting degree. In the breeding and post-breeding seasons, vultures that moved more encountered more conspecifics in flight and while feeding, but this relationship did not hold in the pre-breeding season.

@@

```{r}
#| eval: false
#| output: false
#| label: fig-s-movement
#| fig-cap: "Relationship between normalized strength and movement"
plot_s_eff_movement # null
```

There was no relationship between movement and strength, either overall or in any season or situation.

```{r}
#| eval: false
#| label: fig-s-roost
#| fig-cap: "Relationship between normalized strength and roost diversification"
plot_s_eff_roost
```

However, individuals with more diversified roosts had lower strength (i.e. they interacted less) overall, considering seasons and situations together.

```{r}
#| eval: false
#| label: fig-s-space
#| fig-cap: "Relationship between normalized strength and space use"
plot_s_eff_situ_space
plot_s_emt_situ_space
```

Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network). They also tended to have a higher co-roosting strength, though the trend was not statistically significant. There was no relationship between seasonal space use and co-feeding strength.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r include = FALSE}
library(tidyverse)
library(here)
library(targets)
library(patchwork)
library(ggeffects)
library(gtsummary)
theme_set(theme_classic())
tar_config_set(store = here::here('_targets'))
#source(here("scripts/04.0_mixedModels.R"))
tar_load(cleaned_r)
tar_load(joined0_r)
tar_load(downsampled_10min_forSocial_r)
tar_load(downsampled_10min_r)
tar_load(cc)
tar_load(linked)
tar_load(deg_mod)
tar_load(deg_z_mod)
tar_load(str_mod)
tar_load(str_z_mod)
tar_load(situcolors)
tar_load(seasoncolors)
tar_load(movement_corr)
tar_load(space_corr)
tar_load(roost_corr)
```

@@

```{r}
pctremovedcleaned <- (joined0_r[1]-cleaned_r[1])/joined0_r[1]
```

@@

```{r}
nvultures <- map_dbl(downsampled_10min_forSocial_r, "vultures")
```

@@

```{r}
nvulturesmov <- map_dbl(downsampled_10min_r, "vultures")
```

After removing days with too few fixes per day, we removed any vultures that did not have at least 30 days of valid movement data remaining in a given season, to ensure that movement estimates and averages were based on a high enough sample size of days to capture the range of the vulture's behavior. Since this was done on a per-season basis, individuals that were removed from one season's analysis due to not having enough valid days could still have movement measures calculated for a different season. Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes. The data used for movement analysis contained between `r min(nvulturesmov)` and `r max(nvulturesmov)` vultures per season (mean = `r mean(nvulturesmov)`; sd = `r sd(nvulturesmov)`).

@@

```{r include = FALSE}
# INSERT VARIABLE CONTRIBUTION PLOTS HERE
```

@@

**Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity. It explained [PERCENT]% of the variation in the PCA. Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall. **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE. It explained [PERCENT]% of the variation in the PCA. Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e. the core area made up a larger proportion of the home range). **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night). It explained [PERCENT]% of the variation in the PCA.

@@

## Social preference (deviations from random sociality)

Over the course of a season, individual vultures interacted with significantly more conspecifics, and had significantly more interactions, than expected by chance (i.e. observed degree and strength values were higher than those derived from the randomized datasets). The magnitude of the deviations from random varied by social situation.

```{r message = FALSE}
linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density(aes(linetype = measure))+
  facet_wrap(~measure*situ, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")
```

@@

@@
